---
title: "00_Config"
author: "Jessica Stephens"
date: "7/21/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Install Packages

In a previous session, you saw how to install additional packages that are not available on CRAN, like [ICPIutilities](https://github.com/ICPI/ICPIutilities/blob/master/README.md) which is package developed for use with the PEPFAR MSD. It can be installed from GitHub (via the `devtools` package). If you don't have it installed yet, you can run the code below to do so.

#if difficulties can download zip file 
```{r install_github}
#install ICPI utilities using devtools
#install.packages("devtools")
#library(devtools)
#devtools::install_github("ICPI/ICPIutilities")
#devtools::install_github("USAID-OHA-SI/glitr")

```

There are other useful packages to improve your productivity

1. here
The package here will help you find your files
#helps find files
```{r echo=FALSE}
#install.packages("here")
```

2. glamr
This package is a sister package of ICPIutilities for working with data from datim and the MER
The main function that we will be introducing today is glamr::folder_setup()
#glitter, visualizations, make it easier to use ggplot with standard themes
#glamr, useful things like folder set ups
```{r install_github}
#devtools::install_github("USAID-OHA-SI/glamr")
```




## Load Packages
`janitor` and `dplyr` version >= 1.0.0
Now we will load the packages we want to use.

```{r load, echo=FALSE}
#install.packages("janitor")
#install.packages("tidyverse")
#install.packages("dplyr") 
#package dplyr included in Tidyverse, no need to download separately, unless it has its own update


```


## Load Packages
Now we will load the packages we want to use.

```{r load, echo=FALSE}
library(tidyverse)
library(readxl)
library(here)
library(ICPIutilities)
library(glitr)
library(glamr)
library(janitor)
library(dplyr)

```

## Reset Folders

```{r load, echo=FALSE}
#check that wd is the project, not a subfolder,etc
getwd()

#make sure folder set up is complete
 glamr::folder_setup()
```
