---
title: "01_Munging" --- BK
author: "Jessica Stephens"
date: "7/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### START AT LINE 400 FOR FULL DATASET

## Importing Data OF 4 COUNTRIES

data is in a microsoft excel workbook
readxl

```{r load, echo=FALSE}

#check wd and file path match

here("ZMB MWI KEN UGA by age group DataList_7_9_2020 2_37_41 PM.xlsx")
getwd()
setwd("C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Treatment_Aging_1yrBands")

#read xlsx - specify path, sheet name to read

df_ken_uga <- read_xlsx("Data/ZMB MWI KEN UGA by age group DataList_7_9_2020 2_37_41 PM.xlsx",
                sheet="KEN and UGA") 


df_zmb_mwi <- read_xlsx("Data/ZMB MWI KEN UGA by age group DataList_7_9_2020 2_37_41 PM.xlsx",
                sheet="ZMB and MWI") 

```

## Check Data Import


```{r load, echo=FALSE}

glimpse(df_ken_uga)
View(df_ken_uga)

```



## Append dataset
We can add both df's together to have one df with all of the observations for the 4 countries as a foundational df
will use bind_rows here in the tidyverse, but union can also be used in base R to see if things are compatible

```{r load, echo=FALSE}

df_combined <- bind_rows(df_ken_uga, df_zmb_mwi)


```

See what that looks like

```{r}
print(df_combined)
view(df_combined)
```








## Clean dataset: Isolate Age from E_Ind and drop unnecessary columns/variables

Goal: the numbers for each one-year age band of people with HIV and of people on ART
# make a new variable with the the values after "Age" and before ";"
```{r load, echo=FALSE}



df_combined_seperate <- df_combined  %>% 
    separate(E_Ind, c("A","B", "C", "D", "E"), sep = "([+;])") %>%#"B" holds the age
    separate(B, c("X", "Y", "Age_num"), sep = "([ ])")%>%
    separate(A, c("X1", "PLVHIV_or_ART", "X2"), sep = "([-])")%>%
    separate("PLVHIV_or_ART", c("PLVHIV_or_ART", "Y1"), sep = "([0])") %>% #drop 0 from PLHIV vs ART
    select(E_Count, PLVHIV_or_ART, Age_num, Time, Value, Rounded) %>% #drop X1, X2, Y, X, C, D, E,Y1
    rename(Country=E_Count)#rename country variable
    

view(df_combined_seperate)


df_combined_seperate$Age_num <-
    as.numeric(as.character(unlist(df_combined_seperate$Age_num)))

view(df_combined_seperate)
glimpse(df_combined_seperate)

### EXPORT

```


#############################################
# Restructure data
##############################################

## add variables for Age Categories for ACTUAL Age

Create categories for PEPFAR age brackets
<1
1-4
5-9
10-14
15-19
20-24
25-29
30-34
35-39
40-44
45-49
50+

As well as older fine age brackets
<1
1-9
10-14
15-19
20-24
25-29
30-34
35-39
40-44
45-49
50+

https://stackoverflow.com/questions/12979456/r-code-to-categorize-age-into-group-bins-breaks 


```{r load, echo=FALSE}
#find max
range(df_combined_seperate$Age_num)
#0-80

library(data.table)

PEPFARagebreaks <- c(0,1,5,10,15,20,25,30,35,40,50,85)
PEPFARagelabels <- c("<01","1-4","5-9","10-14","15-19","20-24","25-29","30-34",
               "35-39","40-49","50+")
PEPFARagebreaks18 <- c(0,1,10,15,20,25,30,35,40,50,85)
PEPFARagelabels18 <- c("<01","01-09","10-14","15-19","20-24","25-29","30-34",
               "35-39","40-49","50+")


df_combined_PEPFARage2019 <- setDT(df_combined_seperate) [ , ACTUAL_Age_cat_Fine := cut(Age_num, 
                                breaks = PEPFARagebreaks, 
                                right = FALSE, 
                                labels = PEPFARagelabels)]
view(df_combined_PEPFARage2019)
glimpse(df_combined_PEPFARage2019)



df_combined_PEPFARage2018 <- setDT(df_combined_PEPFARage2019) [ , ACTUAL_Age_cat_SemiFine := cut(Age_num, 
                                breaks = PEPFARagebreaks18, 
                                right = FALSE, 
                                labels = PEPFARagelabels18)]


view(df_combined_PEPFARage2018)

glimpse(df_combined_PEPFARage2018)

names(df_combined_PEPFARage2018)
#df_combined_PEPFARagegroups <- df_combined_PEPFARage2018 %>%
 # select(!c(agegroups, agegroups18, agegroups19))
#names(df_combined_PEPFARagegroups)
#view(df_combined_PEPFARagegroups)
```




# Figure out predicted ages 
#ex, 2017 ages if 2016 aged w/o other loss/gains


```{r load, echo=FALSE}

df_combined_PREDICTagetime <- df_combined_PEPFARage2018 %>%
 mutate(Age_predict=Age_num+1, Time_predict=Time+1)

view(df_combined_PREDICTagetime)



df_combined_PREDICTagegroups <- setDT(df_combined_PREDICTagetime) [ , PREDICT_Age_cat_Fine := cut(Age_predict, 
                                breaks = PEPFARagebreaks, 
                                right = FALSE, 
                                labels = PEPFARagelabels)]

view(df_combined_PREDICTagegroups)
glimpse(df_combined_PREDICTagegroups)



df_combined_PREDICTagegroups <- setDT(df_combined_PREDICTagegroups) [ , PREDICT_Age_cat_SemiFine := cut(Age_predict, 
                                breaks = PEPFARagebreaks18, 
                                right = FALSE, 
                                labels = PEPFARagelabels18)]


view(df_combined_PREDICTagegroups)
glimpse(df_combined_PREDICTagegroups)
names(df_combined_PREDICTagegroups)




```






## Group by Age Groups and Country 
#actual and predicted separately
#merge actual and predicted

```{r load, echo=FALSE}

#only for semi fine brackets at this time

df_long_Grouped_Actual_SemiFine<- df_combined_PREDICTagegroups %>%
  group_by(Country, PLVHIV_or_ART, Time, ACTUAL_Age_cat_SemiFine) %>%
  summarise(Total_Value_Actual=sum(Value))

glimpse(df_long_Grouped_Actual_SemiFine)
view(df_long_Grouped_Actual_SemiFine)


df_long_Grouped_Predict_SemiFine<- df_combined_PREDICTagegroups %>%
  group_by(Country, PLVHIV_or_ART, Time_predict, PREDICT_Age_cat_SemiFine) %>%
  summarise(Total_Value_Predict=sum(Value))

glimpse(df_long_Grouped_Predict_SemiFine)
view(df_long_Grouped_Predict_SemiFine)


#rename predict variables to match actual so they can merge
df_long_Grouped_Predict_SemiFine_vars<- df_long_Grouped_Predict_SemiFine %>%
  rename(Time=Time_predict, ACTUAL_Age_cat_SemiFine=PREDICT_Age_cat_SemiFine)

glimpse(df_long_Grouped_Predict_SemiFine_vars)
view(df_long_Grouped_Predict_SemiFine_vars)

df_long_Grouped_SemiFine <- df_long_Grouped_Actual_SemiFine %>% 
  left_join(df_long_Grouped_Predict_SemiFine_vars, by=c("Time"="Time", "ACTUAL_Age_cat_SemiFine"="ACTUAL_Age_cat_SemiFine","Country"="Country", "PLVHIV_or_ART"="PLVHIV_or_ART"))

view(df_long_Grouped_SemiFine)





```










## look at difference between actual ages - aging in/out actual loss or growth 
#ex actual 2019 / actual 2018

##calc predicted % difference
#predicted 2019/actual 2018

##Compare actual to predicted - expected impact based on aging


```{r load, echo=FALSE}

## BK's solution
## Note: Make sure to install `janitor` and `dplyr` version >= 1.0.0
## This version calculates the perc for all years
df_perc_diff<- df_long_Grouped_SemiFine %>%
  janitor::clean_names() %>% 
  arrange(actual_age_cat_semi_fine) %>% #making year values sequential 
  group_by(country, actual_age_cat_semi_fine) %>% #dont want to confuse the countries or hiv/art obs - keep lag/lead in these groups
  mutate(
    total_prev_value_actual = lag(total_value_actual),
  #   perc_diff_actual = ifelse(
  #     !is.na(total_value_actual) & !is.na(total_prev_value_actual), #ignore the values that are empty (na)
  #     round((total_value_actual - total_prev_value_actual) / total_prev_value_actual * 100), #this is rounded to the whole number, but can put a "," after "100" (function round)
  #     NA
  #   ),
     perc_diff_actual2 = ifelse(
    !is.na(total_value_actual) & !is.na(total_prev_value_actual), #ignore the values that are empty (na)
    total_value_actual / total_prev_value_actual*100, #this is rounded to the whole number, but can put a "," after "100" (function round)
    NA
    ),
    # perc_diff_predict = ifelse(
    #   !is.na(total_value_predict) & !is.na(lag(total_value_actual)),
    #   round((total_value_predict - total_prev_value_actual) / total_prev_value_actual * 100),
    #   NA
    # ),
    perc_diff_predict2 = ifelse(
      !is.na(total_value_predict) & !is.na(lag(total_value_actual)),
      total_value_predict / total_prev_value_actual * 100,
      NA
    ),
 #   perc_diff_Compare= perc_diff_actual-perc_diff_predict
    perc_diff_Compare2= (perc_diff_actual2/perc_diff_predict2)*100
  ) %>% 
  ungroup() %>% 
  relocate(total_prev_value_actual, .after = total_value_actual)
  View(df_perc_diff)

## This still works
## Transform from long to wide
## Pay attention to the use of `lag` for `Perc_diff_predict` variables
# df_wide_PEPFARagegroups <- df_long_Grouped_SemiFine %>%
#   gather(key="counttype", value= "countvalue", -c(1:4)) %>%
#   spread(Time, countvalue) %>%
#   mutate(
#     counttype=str_remove(counttype,"Total_Value_"),
#     Perc_diff_actual= ifelse(
#       counttype == "Actual", 
#       round((`2019` - `2018`) / `2018` * 100), 
#       NA
#     ), 
#     Perc_diff_perdict = ifelse(
#       counttype == "Predict",
#       round((`2019` - lag(`2018`)) / `2018` * 100),
#       NA
#     )
#   ) %>% 
#   View()

#ex actual 2019 / actual 2018
#predicted 2019/actual 2018
#not sure if this works

# df_wide_PEPFARagegroups <- df_combined_PEPFARage2018 %>%
#   pivot_wider(
#       names_from=Time,
#       names_glue="ACTUAL_Value_{Time}",
#       values_from=Value
#       )

#view(df_wide_PEPFARagegroups)

  

```










#filter ART, time and select relevant variables


```{r load, echo=FALSE}
spectrum_clean<-df_perc_diff%>%
  filter(time %in% c('2018', '2019')) %>%
  select(country, time, actual_age_cat_semi_fine, perc_diff_predict2)%>%
  view()
view(spectrum_clean)

```















##########################################################
## Repeat for new full data set for all PEPFAR OUs
##########################################################





---
title: "01_Munging_HIV2020Estimates_ART by single year_8Sept2020_norow"
author: "Jessica Stephens"
date: "9/11/2020"
output: html_document
---



## Importing Data

data is in a microsoft excel workbook
readxl

```{r load, echo=FALSE}

#check wd and file path match


setwd("C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Treatment_Aging_1yrBands")

#read xlsx - specify path, sheet name to read

df <- read_xlsx("Data/HIV2020Estimates_ART by single year_8Sept2020_norow.xlsx",
                sheet="DataList") 


```


## Check Data Import


```{r load, echo=FALSE}

glimpse(df)
View(df)

```




## Clean dataset: Isolate Age from E_Ind and drop unnecessary columns/variables

Goal: the numbers for each one-year age band of people with HIV and of people on ART
# make a new variable with the the values after "Age" and before ";"
```{r load, echo=FALSE}



df_full_seperate <- df %>% 
    separate(E_Ind, c("A","B", "C", "D", "E"), sep = "([+;])") %>%#"B" holds the age
    separate(B, c("X", "Y", "Age_num"), sep = "([ ])")%>%
    separate(A, c("X1", "PLVHIV_or_ART", "X2"), sep = "([-])")%>%
    separate("PLVHIV_or_ART", c("PLVHIV_or_ART", "Y1"), sep = "([0])") %>% #drop 0 from PLHIV vs ART
    rename(Country=E_Count, sex=D, sex2=E)  %>%#rename country variable and vars with sex
    mutate(malefemale=ifelse(is.na(sex2), 0, 1)) %>%
    mutate(sex=ifelse(malefemale==1, "MaleFemale", sex)) %>% # not working to change sex here %>%
    mutate(Country=recode(Country, "Cote dIvoire"="Cote d'Ivoire", 
                          "Lao People Democratic Republic"="Laos",
                          "United Republic of Tanzania"= "Tanzania",
                          "Viet Nam"="Vietnam"))%>%
    select(Country, Age_num, sex, Time, Value) %>% #drop X1, X2, Y, X, C,Y1, sex2 Rounded
    view() %>%
    glimpse()
  


#unique(c(as.character(df_full_seperate$sex)),as.character(df_full_seperate$sex2))

#   Country PLVHIV_or_ART            Age_num Male    Female  Time Value Rounded
#   <chr>   <chr>                    <chr>   <chr>   <chr>  <dbl> <dbl>   <dbl>
# 1 Angola  " Number on ART by age " 0       " Male" Female  2010   163     163


df_full_seperate$Age_num <-
    as.numeric(as.character(unlist(df_full_seperate$Age_num)))

view(df_full_seperate)
glimpse(df_full_seperate)

```



#############################################
# Restructure data
##############################################

## add variables for Age Categories for ACTUAL Age

Create categories for PEPFAR age brackets
<1
1-4
5-9
10-14
15-19
20-24
25-29
30-34
35-39
40-44
45-49
50+

As well as older fine age brackets
<1
1-9
10-14
15-19
20-24
25-29
30-34
35-39
40-44
45-49
50+

https://stackoverflow.com/questions/12979456/r-code-to-categorize-age-into-group-bins-breaks 


```{r load, echo=FALSE}
#find max
range(df_full_seperate$Age_num)
#0-80

library(data.table)

# PEPFARagebreaks <- c(0,1,5,10,15,20,25,30,35,40,50,85)
# PEPFARagelabels <- c("<01","1-4","5-9","10-14","15-19","20-24","25-29","30-34",
#                "35-39","40-49","50+")
PEPFARagebreaks18 <- c(0,1,10,15,20,25,30,35,40,50,85)
PEPFARagelabels18 <- c("<01","01--09","10--14","15-19","20-24","25-29","30-34",
               "35-39","40-49","50+")


# df_combined_PEPFARage2019 <- setDT(df_combined_seperate) [ , ACTUAL_Age_cat_Fine := cut(Age_num, 
#                                 breaks = PEPFARagebreaks, 
#                                 right = FALSE, 
#                                 labels = PEPFARagelabels)]
# view(df_combined_PEPFARage2019)
# glimpse(df_combined_PEPFARage2019)



df_full_PEPFARage2018 <- setDT(df_full_seperate) [ , ACTUAL_Age_cat_SemiFine := cut(Age_num, 
                                breaks = PEPFARagebreaks18, 
                                right = FALSE, 
                                labels = PEPFARagelabels18)]


view(df_full_PEPFARage2018)

glimpse(df_full_PEPFARage2018)

names(df_full_PEPFARage2018)

```




# Figure out predicted ages 
#ex, 2017 ages if 2016 aged w/o other loss/gains


```{r load, echo=FALSE}

df_full_PREDICTagetime <- df_full_PEPFARage2018 %>%
 mutate(Age_predict=Age_num+1, Time_predict=Time+1)

view(df_full_PREDICTagetime)


# df_combined_PREDICTagegroups <- setDT(df_combined_PREDICTagetime) [ , PREDICT_Age_cat_Fine := cut(Age_predict, 
#                                 breaks = PEPFARagebreaks, 
#                                 right = FALSE, 
#                                 labels = PEPFARagelabels)]
# 
# view(df_combined_PREDICTagegroups)
# glimpse(df_combined_PREDICTagegroups)


df_full_PREDICTagegroups <- setDT(df_full_PREDICTagetime) [ , PREDICT_Age_cat_SemiFine := cut(Age_predict, 
                                breaks = PEPFARagebreaks18, 
                                right = FALSE, 
                                labels = PEPFARagelabels18)]


view(df_full_PREDICTagegroups)
glimpse(df_full_PREDICTagegroups)
#Rows: 119,280
#Columns: 11
names(df_full_PREDICTagegroups)




```



##############################################################################
Multiple data formats
1) long with disaggregate (<15 no disaggregate)
2) long without sex disaggregate
3) long with all data (male, female and malefemale) - can be disaggregated later

## Start with long non-disaggregated because it has all age groups "MaleFemale"

#### SKIP BELOW FOR SEX DISAGGREGATE DATA OR ALL THE WAY TO BOTTOM TOTAL DATASET (MALEFEMALE, MALE, FEMALE)


###########################################################
######    NO DISAGGREGATE (format 1) ######
###########################################################

```{r load, echo=FALSE}

#2
  #drop the malefemale==0 and sex

df_long_total <- df_full_PREDICTagegroups %>%
    subset(sex="MaleFemale") %>%
    select(!sex) %>%
    glimpse()


# df_long_total <- df_full_PREDICTagegroups %>%
#     subset(malefemale==1) %>%
#     select(!malefemale) %>%
#     select(!sex) %>%
#     glimpse()
#   
view(df_long_total)
glimpse(df_long_total)

#Rows: 45,360
#Columns: 9
# 
# write_csv(df_long_total,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/Spectrum_long_20200914.csv", na="")
```



## Group by Age Groups and Country 
#actual and predicted separately
#merge actual and predicted

```{r load, echo=FALSE}

#insert the df_long_total

#actual
df_long_Grouped_Actual_SemiFine<- df_long_total %>%
  group_by(Country, Time, ACTUAL_Age_cat_SemiFine) %>%
  summarise(Total_Value_Actual=sum(Value))

glimpse(df_long_Grouped_Actual_SemiFine)
view(df_long_Grouped_Actual_SemiFine)

#predict
df_long_Grouped_Predict_SemiFine<- df_long_total %>%
  group_by(Country, Time_predict, PREDICT_Age_cat_SemiFine) %>%
  summarise(Total_Value_Predict=sum(Value))

glimpse(df_long_Grouped_Predict_SemiFine)
view(df_long_Grouped_Predict_SemiFine)


#rename predict variables to match actual so they can merge
df_long_Grouped_Predict_SemiFine_vars<- df_long_Grouped_Predict_SemiFine %>%
  rename(Time=Time_predict, ACTUAL_Age_cat_SemiFine=PREDICT_Age_cat_SemiFine)

glimpse(df_long_Grouped_Predict_SemiFine_vars)
view(df_long_Grouped_Predict_SemiFine_vars)

#merge back together
df_long_Grouped_SemiFine <- df_long_Grouped_Actual_SemiFine %>% 
  left_join(df_long_Grouped_Predict_SemiFine_vars, by=c("Time"="Time", "ACTUAL_Age_cat_SemiFine"="ACTUAL_Age_cat_SemiFine","Country"="Country"))

view(df_long_Grouped_SemiFine)





```






## look at difference between actual ages - aging in/out actual loss or growth 
#ex actual 2019 / actual 2018

##calc predicted % difference
#predicted 2019/actual 2018

##Compare actual to predicted - expected impact based on aging


```{r load, echo=FALSE}

## BK's solution
## Note: Make sure to install `janitor` and `dplyr` version >= 1.0.0
## This version calculates the perc for all years
df_perc<- df_long_Grouped_SemiFine %>%
  janitor::clean_names() %>% 
  arrange(actual_age_cat_semi_fine) %>% #making year values sequential 
  group_by(country, actual_age_cat_semi_fine) %>% #dont want to confuse the countries or hiv/art obs - keep lag/lead in these groups
  mutate( total_prev_value_actual = lag(total_value_actual),
    
    perc_diff = ifelse(
    !is.na(total_value_actual) & !is.na(total_prev_value_actual), #ignore the values that are empty (na)
    total_value_actual / total_prev_value_actual*100, 
    NA
    ),
    
    perc_of_lastyr = ifelse(
      !is.na(total_value_predict) & !is.na(lag(total_value_actual)),
      total_value_predict / total_prev_value_actual * 100,
      NA
#  ),
 #       perc_diff_Compare2= (perc_diff/perc_of_lastyr)*100
 )) %>% 
  ungroup() %>% 
  relocate(total_prev_value_actual, .after = total_value_actual)
  
View(df_perc)
  


  

```



#filter ART, time and select relevant variables


```{r load, echo=FALSE}
spectrum_clean_all<-df_perc%>%
  filter(time %in% c( '2019')) %>%
  select(country, time, actual_age_cat_semi_fine,   perc_of_lastyr)%>%
    rename(year=time, age_group=actual_age_cat_semi_fine)%>%
    mutate_if(is.factor, as.character)%>%
  glimpse()

view(spectrum_clean_all)
write_csv(spectrum_clean_all,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/spectrum_clean_all_20200916.csv", na="")

```










###########################################################
######    SEX DISAGGREGATE (format 2) ######
###########################################################


Multiple data formats
1) long with disaggregate
2) long without sex disaggregate
3) long with all data (male, female and malefemale) - can be disaggregated later
```{r load, echo=FALSE}


#1
  #drop the malefemale==1
df_long_sexdisagg <- df_full_PREDICTagegroups %>%
    subset(sex!=2) %>%
    view()

view(df_long_sexdisagg)
glimpse(df_long_sexdisagg)
#Rows: 73,920
#Columns: 10

# write_csv(df_long_sexdisagg,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/Spectrum_long_sexdisagg_20200914.csv", na="")

```





## Group by Age Groups and Country and sex
#actual and predicted separately
#merge actual and predicted

```{r load, echo=FALSE}

#insert the df_long_sexdisagg

#actual
df_long_Grouped_Actual_SemiFine_disagg<- df_long_sexdisagg %>%
  group_by(Country, sex, Time, ACTUAL_Age_cat_SemiFine) %>%
  summarise(Total_Value_Actual=sum(Value))

glimpse(df_long_Grouped_Actual_SemiFine_disagg)
view(df_long_Grouped_Actual_SemiFine_disagg)

#predict
df_long_Grouped_Predict_SemiFine_disagg<- df_long_sexdisagg %>%
  group_by(Country, sex, Time_predict, PREDICT_Age_cat_SemiFine) %>%
  summarise(Total_Value_Predict=sum(Value))

glimpse(df_long_Grouped_Predict_SemiFine_disagg)
view(df_long_Grouped_Predict_SemiFine_disagg)


#rename predict variables to match actual so they can merge
df_long_Grouped_Predict_SemiFine_vars_disagg<- df_long_Grouped_Predict_SemiFine_disagg %>%
  rename(Time=Time_predict, ACTUAL_Age_cat_SemiFine=PREDICT_Age_cat_SemiFine)

glimpse(df_long_Grouped_Predict_SemiFine_vars_disagg)
view(df_long_Grouped_Predict_SemiFine_vars_disagg)

#merge back together
df_long_Grouped_SemiFine_disagg <- df_long_Grouped_Actual_SemiFine_disagg %>% 
  left_join(df_long_Grouped_Predict_SemiFine_vars_disagg, by=c("Time"="Time", "ACTUAL_Age_cat_SemiFine"="ACTUAL_Age_cat_SemiFine","Country"="Country", "sex"="sex"))

view(df_long_Grouped_SemiFine_disagg)





```






## look at difference between actual ages - aging in/out actual loss or growth 
#ex actual 2019 / actual 2018

##calc predicted % difference
#predicted 2019/actual 2018

##Compare actual to predicted - expected impact based on aging


```{r load, echo=FALSE}

## BK's solution
## Note: Make sure to install `janitor` and `dplyr` version >= 1.0.0
## This version calculates the perc for all years
df_perc_disagg<- df_long_Grouped_SemiFine_disagg %>%
  janitor::clean_names() %>% 
  arrange( actual_age_cat_semi_fine) %>% #making year values sequential 
  group_by(country, actual_age_cat_semi_fine) %>% #dont want to confuse the countries or hiv/art obs - keep lag/lead in these groups
  mutate( total_prev_value_actual = lag(total_value_actual),
    
    perc_diff = ifelse(
    !is.na(total_value_actual) & !is.na(total_prev_value_actual), #ignore the values that are empty (na)
    total_value_actual / total_prev_value_actual*100, 
    NA
    ),
    
    perc_of_lastyr = ifelse(
      !is.na(total_value_predict) & !is.na(lag(total_value_actual)),
      total_value_predict / total_prev_value_actual * 100,
      NA
#  ),
 #       perc_diff_Compare2= (perc_diff/perc_of_lastyr)*100
 )) %>% 
  ungroup() %>% 
  relocate(total_prev_value_actual, .after = total_value_actual)
  
View(df_perc_disagg)
  


  

```



#filter ART, time and select relevant variables


```{r load, echo=FALSE}
spectrum_clean_disagg<-df_perc_disagg%>%
  filter(time %in% c('2019')) %>%
  select(country, time, sex, actual_age_cat_semi_fine, perc_of_lastyr)%>%
  rename(year=time, age_group=actual_age_cat_semi_fine)%>%
  mutate_if(is.factor, as.character)%>%
  mutate(sex=ifelse(sex==" Female", "Female", "Male")) %>%
  glimpse()

view(spectrum_clean_disagg)
glimpse(spectrum_clean_disagg)

write_csv(spectrum_clean_disagg,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/spectrum_clean_sexdisagg_20200916.csv", na = "")

```

















###########################################################
######    ALL DATA (format 3) ######
###########################################################


Multiple data formats
1) long with disaggregate
2) long without sex disaggregate
3) long with all data (male, female and malefemale) - can be disaggregated later


## Group by Age Groups and Country and sex
#actual and predicted separately
#merge actual and predicted

```{r load, echo=FALSE}

#insert the df_long_sexdisagg

#actual
df_long_Grouped_Actual_SemiFine_disagg<- df_full_PREDICTagegroups %>%
  group_by(Country, sex, Time, ACTUAL_Age_cat_SemiFine) %>%
  summarise(Total_Value_Actual=sum(Value))

glimpse(df_long_Grouped_Actual_SemiFine_disagg)
view(df_long_Grouped_Actual_SemiFine_disagg)

#predict
df_long_Grouped_Predict_SemiFine_disagg<- df_full_PREDICTagegroups %>%
  group_by(Country, sex, Time_predict, PREDICT_Age_cat_SemiFine) %>%
  summarise(Total_Value_Predict=sum(Value))

glimpse(df_long_Grouped_Predict_SemiFine_disagg)
view(df_long_Grouped_Predict_SemiFine_disagg)


#rename predict variables to match actual so they can merge
df_long_Grouped_Predict_SemiFine_vars_disagg<- df_long_Grouped_Predict_SemiFine_disagg %>%
  rename(Time=Time_predict, ACTUAL_Age_cat_SemiFine=PREDICT_Age_cat_SemiFine)

glimpse(df_long_Grouped_Predict_SemiFine_vars_disagg)
view(df_long_Grouped_Predict_SemiFine_vars_disagg)

#merge back together
df_long_Grouped_SemiFine_disagg <- df_long_Grouped_Actual_SemiFine_disagg %>% 
  left_join(df_long_Grouped_Predict_SemiFine_vars_disagg, by=c("Time"="Time", "ACTUAL_Age_cat_SemiFine"="ACTUAL_Age_cat_SemiFine","Country"="Country", "sex"="sex"))

view(df_long_Grouped_SemiFine_disagg)
# 
# write_csv(df_long_Grouped_SemiFine_disagg,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/Spectrum_long_grouped_20200914.csv", na="")



```






## look at difference between actual ages - aging in/out actual loss or growth 
#ex actual 2019 / actual 2018

##calc predicted % difference
#predicted 2019/actual 2018

##Compare actual to predicted - expected impact based on aging


```{r load, echo=FALSE}

## BK's solution
## Note: Make sure to install `janitor` and `dplyr` version >= 1.0.0
## This version calculates the perc for all years
df_perc_disagg<- df_long_Grouped_SemiFine_disagg %>%
  janitor::clean_names() %>% 
  arrange( actual_age_cat_semi_fine) %>% #making year values sequential 
  group_by(country, actual_age_cat_semi_fine) %>% #dont want to confuse the countries or hiv/art obs - keep lag/lead in these groups
  mutate( total_prev_value_actual = lag(total_value_actual),
    
    perc_diff = ifelse(
    !is.na(total_value_actual) & !is.na(total_prev_value_actual), #ignore the values that are empty (na)
    total_value_actual / total_prev_value_actual*100, 
    NA
    ),
    
    perc_of_lastyr = ifelse(
      !is.na(total_value_predict) & !is.na(lag(total_value_actual)),
      total_value_predict / total_prev_value_actual * 100,
      NA
#  ),
 #       perc_diff_Compare2= (perc_diff/perc_of_lastyr)*100
 )) %>% 
  ungroup() %>% 
  relocate(total_prev_value_actual, .after = total_value_actual)
  
View(df_perc_disagg)
  


  

```



#filter ART, time and select relevant variables


```{r load, echo=FALSE}
spectrum_clean_all_disagg<-df_perc_disagg%>%
  filter(time %in% c('2019')) %>%
  select(country, time, sex, actual_age_cat_semi_fine, perc_of_lastyr)%>%
  rename(year=time, age_group=actual_age_cat_semi_fine)%>%
  mutate_if(is.factor, as.character)%>%
  mutate(sex=ifelse(sex==" Female", "Female", sex)) %>%
  mutate(sex=ifelse(sex==" Male", "Male", sex)) %>%
  glimpse()

view(spectrum_clean_all_disagg)
glimpse(spectrum_clean_all_disagg)

write_csv(spectrum_clean_all_disagg,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/spectrum_clean_all_total_20200916.csv", na="")

```






























#################################   ARCHIVE CODE ##############################


<!-- # Figure out predicted ages  -->
<!-- #ex, 2017 ages if 2016 aged w/o other loss/gains -->


<!-- ```{r load, echo=FALSE} -->



<!-- PREDICTagebreaks <- c(0,4,9,14,19,24,29,34,39,44,49,85) -->
<!-- PREDICTagelabels <- c("0-3","4-8","9-13","14-18","19-23","24-28","29-33", -->
<!--                "34-38","39-43","44-48","49+") -->

<!-- PREDICTagebreaks18 <- c(0,9,14,19,24,29,34,39,44,49,85) -->
<!-- PREDICTagelabels18 <- c("0-8","9-13","14-18","19-23","24-28","29-33", -->
<!--                "34-38","39-43","44-48","49+") -->

<!-- df_wide_PREDICTagegroupsFINE <- setDT(df_combined_PEPFARagegroups) [ , PREDICT_Age_cat_Fine := cut(Age_num,  -->
<!--                                 breaks = PREDICTagebreaks,  -->
<!--                                 right = FALSE,  -->
<!--                                 labels = PREDICTagelabels)] -->
<!-- view(df_wide_PREDICTagegroupsFINE) -->
<!-- glimpse(df_wide_PREDICTagegroupsFINE) -->

<!-- df_wide_PREDICTagegroupsALL <- setDT(df_wide_PREDICTagegroupsFINE) [ , PREDICT_Age_cat_SemiFine := cut(Age_num,  -->
<!--                                 breaks = PREDICTagebreaks18,  -->
<!--                                 right = FALSE,  -->
<!--                                 labels = PREDICTagelabels18)] -->
<!-- view(df_wide_PREDICTagegroupsALL) -->
<!-- glimpse(df_wide_PREDICTagegroupsALL) -->

<!-- ``` -->



<!-- ## Figure out predicted ages  -->
<!-- #ex, 2017 ages if 2016 aged w/o other loss/gains -->


<!-- ```{r load, echo=FALSE} -->


<!-- for each year and actual age cat -->
<!--   do value of previous year and  -->


<!-- PRED_Value_2019(ACTUAL_Age_cat_Fine) = ACTUAL_Value_2018 (PRED_age_bands) -->


<!-- #loop over age bands -->

<!-- #long -->

<!-- for(i in Value){ -->
<!--   Pred_Value <- when Time=Time-1 & PRED_Age_cat_SemiFine=0 -->



<!--  # wide??? -->

<!--   sum(ACTUAL_Value_2018) if PRED_age_bands =  -->
<!-- } -->

<!-- PRED_Value_2019_1-9  <- ACTUAL_Value_2018_0-8 -->

<!-- PRED_Value_2019_1-9  <- ACTUAL_Value_2018_0-8 -->


<!-- ``` -->



