---
title: "02_MSD Munging"
author: "Jessica Stephens"
date: "8/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



###################### INPORT AND MUNGE MSD #############################


## Merge Tx_Curr and TX_new
## Downloaded from Panorama, MSD OU by IM for 2019

```{r load, echo=FALSE}

#msd <- read_msd("C:/Users/jesse/OneDrive/Documents/Work Material/USAID/MER/FY19/MER_Structured_Datasets_OU_IM_FY17-20_20191220_v2_1_MER FY19Q4 Post Clean/MER_Structured_Datasets_OU_IM_FY17-20_20191220_v2_1.txt", save_rds=FALSE)

msd <- read_msd("../Data/MER_Structured_Datasets_OU_IM_FY17-20_20191220_v2_1.txt", save_rds=FALSE)


glimpse(msd)
  

```



############## MUNGE MSD (DO NOT INCLUDE SEX, USING TOTALS) ################


```{r load, echo=FALSE}


msd_format <-
  msd %>%
  filter(
    #operatingunit %in% c("Uganda", "Kenya", "Malawi", "Zambia"),
         indicator %in% c("TX_CURR", "TX_NEW"),
         fiscal_year %in% c('2018', '2019'),
         trendsfine!="Retired Age Band",
         trendsfine!="Coarse",
         trendsfine!="Unknown Age",
         standardizeddisaggregate== "Age/Sex/HIVStatus") %>% #double check
  select(operatingunit, indicator, trendsfine, fiscal_year, cumulative, sex) %>%
  arrange(operatingunit,indicator, trendsfine, fiscal_year)  %>%
#  view()
  group_by(operatingunit,indicator, trendsfine, fiscal_year) %>%
  summarize(
    cumulative_total=sum(cumulative,na.rm=TRUE)) %>%
 #   n=n()) #will show how many rows in grouping --- also can use count() function alone instead of summ with n=n()
   view()

view(msd_format)

#  gather(key="cum", value="total", -c(1:4)) %>%
#  gather(key="cum", value="total", -operatingunit,-indicator, -trendsfine, -fiscal_year) %>%

```




#long to wide so that tx_Curr and TX_new are on same row for country and year and age
#change var names so mergable with spectrum data


```{r load, echo=FALSE}

 msd_format_wide <- msd_format %>%
  #long to wide so that tx_Curr and TX_new are on same row for country and year and age
  spread(indicator, cumulative_total) %>%
  rename(country=operatingunit, year=fiscal_year, age_group=trendsfine)%>%  
  pivot_wider(
     names_from=year,
     names_glue="{.value}_{year}",
     values_from=c(TX_CURR, TX_NEW)
  )  %>%
  #change var names so mergable with spectrum data
    mutate_if(~ is.integer(.), as.double)%>%
  mutate(age_group=recode(age_group, "<1"="<01",
                          "01-09"="01--09",
                          "10-14"="10--14")) %>%
  select(!TX_NEW_2018)%>%
  glimpse()
  
  View(msd_format_wide)
  glimpse(msd_format_wide)

```






#######################  merge with spectrum data #######################  

#### START WITH FORMAT 1, NO DISAGGRGATE

<!-- spectrum_clean_all<-read_csv("C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/spectrum_clean_all_20200916.csv") -->

```{r load, echo=FALSE}

spectrum_clean_all <- read_csv("../Data/Output/spectrum_clean_all_20200916.csv")
   
glimpse(spectrum_clean_all)

#also could have subset from larger dataset
# spectrum_clean_all2 <- spectrum_clean_all_disagg %>%
#     subset(sex="MaleFemale) %>%
#     select(!sex) %>%
#     view()

df_msd_spectrum <- spectrum_clean_all  %>% 
  left_join(msd_format_wide, 
            by=c("country"="country", "age_group"="age_group"), 
            copy = TRUE)  %>% 
  select(!year)

view(df_msd_spectrum)


```


#calclated variables


```{r load, echo=FALSE}


df_calulcated_vars <-df_msd_spectrum %>% 
  rowwise() %>% 
  mutate(
         ## Calc predicted remaining in 2019
        # prev year tx_curr*prediction % difference
        TX_CURR_2019_aging_adj_baseline=round(TX_CURR_2018 * (perc_of_lastyr/100) ),
         ##Calc Expected FY19 CURR accounting for aging in/out
          #current FY tx_new*predicted remaning in 2019 (last calculation)
         TX_CURR_2019_aging_adj_final= sum(TX_CURR_2019_aging_adj_baseline, 
                                           TX_NEW_2019, na.rm = TRUE),
         retention_proxy_aging_adj=((TX_CURR_2019/TX_CURR_2019_aging_adj_final)*100),###redone
         retention_proxy_pepfar=((TX_CURR_2019/(TX_CURR_2018+TX_NEW_2019))*100), #redone
         retention_highest=(ifelse((100-retention_proxy_aging_adj) < (100-retention_proxy_pepfar),"Aging Adjusted Proxy", "PEPFAR Proxy")))

view(df_calulcated_vars)

glimpse(df_calulcated_vars)
  
write_csv(df_calulcated_vars,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/Spectrum_Retention_20200914.csv", na="")



#split save is icpi utilities package for data export
# #export one csv for each OU for TX_NEW
#   df_ou_im %>%
#     filter(indicator == "TX_NEW",
#            standardizeddisaggregate == "Total Numerator") %>%
#     split_save(operatingunit, "~/Data/FY18Q3", "FY18Q3_TX")
# 
# #export one csv for each funding agency with today's date
#   df_ou_im %>%
#     filter(indicator %in% c("HTS_TST", "HTS_TST_POS", "TX_NEW"),
#            standardizeddisaggregate == "Total Numerator") %>%
#     split_save(fundingagency, "~/CountryFiles", "FY18Q3_Cascade")
```





```{r load, echo=FALSE}

# 
# df<-read_csv("C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/Spectrum_Retention_20200911.csv", sheet="Spectrum_Retention_20200911")

# Spectrum_Retention_20200911 <- read.csv("~/Work Material/USAID/Treatment/Data/Output/Spectrum_Retention_20200911.csv", stringsAsFactors=FALSE)
# View(Spectrum_Retention_20200911)
# glimpse(Spectrum_Retention_20200911)
# Spectrum_Retention_20200911 <- Spectrum_Retention_20200911 %>%
# write_csv(df_calulcated_vars,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/Spectrum_Retention_20200911.csv", na="")


```










############## MUNGE MSD TO INCLUDE SEX ################



```{r load, echo=FALSE}


msd_format <-
  msd %>%
  filter(
    #operatingunit %in% c("Uganda", "Kenya", "Malawi", "Zambia"),
         indicator %in% c("TX_CURR", "TX_NEW"),
         fiscal_year %in% c('2018', '2019'),
         trendsfine!="Retired Age Band",
         trendsfine!="Coarse",
         trendsfine!="Unknown Age",
         standardizeddisaggregate== "Age/Sex/HIVStatus", 
         sex!="Unknown Sex") %>% 
  select(operatingunit, indicator, trendsfine, fiscal_year, cumulative, sex) %>%
  arrange(operatingunit,indicator, trendsfine, fiscal_year, sex)  %>%
#  view()
  group_by(operatingunit,indicator, trendsfine, fiscal_year, sex) %>%
  summarize(
    cumulative=sum(cumulative,na.rm=TRUE)) %>%  

 #   n=n()) #will show how many rows in grouping --- also can use count() function alone instead of summ with n=n()
   view()
view(msd_format)

#  gather(key="cum", value="total", -c(1:4)) %>%
#  gather(key="cum", value="total", -operatingunit,-indicator, -trendsfine, -fiscal_year) %>%

```




#long to wide so that tx_Curr and TX_new are on same row for country and year and age
#change var names so mergable with spectrum data


```{r load, echo=FALSE}

 msd_format_wide <- msd_format %>%
  #long to wide so that tx_Curr and TX_new are on same row for country and year and age
  spread(indicator, cumulative) %>%
  rename(country=operatingunit, year=fiscal_year, age_group=trendsfine)%>%  
  pivot_wider(
     names_from=year,
     names_glue="{.value}_{year}",
     values_from=c(TX_CURR, TX_NEW)
  )  %>%
  #change var names so mergable with spectrum data
  mutate_if(~ is.integer(.), as.double)%>%
  mutate(age_group=recode(age_group, "<1"="<01",
                          "01-09"="01--09",
                          "10-14"="10--14")) %>%
  select(!TX_NEW_2018)%>%
  glimpse()
  
  View(msd_format_wide)
  glimpse(msd_format_wide)

```



#######################  merge with spectrum data #######################  

### MERGE WITH SPECTRUM DATA FORMAT 3 - SEX DISAGGREAGATED 

<!-- spectrum_clean_all_disagg<-read_csv("C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/spectrum_clean_all_total_20200916.csv") -->

```{r load, echo=FALSE}
  glimpse(spectrum_clean_all)

spectrum_clean_disag <- spectrum_clean_all_disagg %>%
    subset(sex!="MaleFemale") %>%
    view()


df_msd_spectrum <-spectrum_clean_disag  %>% 
  left_join(msd_format_wide, by=c("country"="country", "age_group"="age_group", "sex"="sex"), `copy` = TRUE)  %>% 
  select(!year)
view(df_msd_spectrum)


```


#calclated variables


```{r load, echo=FALSE}

df_calulcated_vars <-df_msd_spectrum %>% 
  mutate(
         ## Calc predicted remaining in 2019
        # prev year tx_curr*prediction % difference
        TX_CURR_2019_aging_adj_baseline=round(TX_CURR_2018*(perc_of_lastyr/100)),
         ##Calc Expected FY19 CURR accounting for aging in/out
          #current FY tx_new*predicted remaning in 2019 (last calculation)
         TX_CURR_2019_aging_adj_final=round(TX_CURR_2019_aging_adj_baseline+TX_NEW_2019),
         retention_proxy_aging_adj=((TX_CURR_2019/TX_CURR_2019_aging_adj_final)*100),###redone
         retention_proxy_pepfar=((TX_CURR_2019/(TX_CURR_2018+TX_NEW_2019))*100), #redone
         retention_highest=(ifelse((100-retention_proxy_aging_adj) < (100-retention_proxy_pepfar),"Aging Adjusted Proxy", "PEPFAR Proxy")))


view(df_calulcated_vars)
glimpse(df_calulcated_vars)
  
write_csv(df_calulcated_vars,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/Spectrum_Retention_disagg_20200916.csv", na="")

#write.csv(df_calulcated_vars,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/Spectrum_Retention_20200911.csv", row.names = FALSE)



### EXAMPLE FOR OTHER EXPORT PACKAGE

#split save is icpi utilities package for data export
# #export one csv for each OU for TX_NEW
#   df_ou_im %>%
#     filter(indicator == "TX_NEW",
#            standardizeddisaggregate == "Total Numerator") %>%
#     split_save(operatingunit, "~/Data/FY18Q3", "FY18Q3_TX")
# 
# #export one csv for each funding agency with today's date
#   df_ou_im %>%
#     filter(indicator %in% c("HTS_TST", "HTS_TST_POS", "TX_NEW"),
#            standardizeddisaggregate == "Total Numerator") %>%
#     split_save(fundingagency, "~/CountryFiles", "FY18Q3_Cascade")
```























<!-- ```{r load, echo=FALSE} -->

<!-- #  -->
<!-- # df<-read_csv("C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/Spectrum_Retention_20200911.csv", sheet="Spectrum_Retention_20200911") -->

<!-- Spectrum_Retention_20200911 <- read.csv("~/Work Material/USAID/Treatment/Data/Output/Spectrum_Retention_20200911.csv", stringsAsFactors=FALSE) -->
<!-- View(Spectrum_Retention_20200911) -->
<!-- glimpse(Spectrum_Retention_20200911) -->
<!-- Spectrum_Retention_20200911 <- Spectrum_Retention_20200911 %>% -->
<!--   # mutate(age_group=ifelse("9-Jan", "1--9", -->
<!--   #                         ifelse("14-Oct", "10--14",age_group))) -->

<!-- write_csv(df_calulcated_vars,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/Spectrum_Retention_20200911.csv", na="") -->


<!-- # malefemale=ifelse(is.na(sex2), 0, 1)) %>% -->
<!-- ``` -->

